---
title: "Git with R"
author: "AY"
date: "`r format(Sys.time(),'%b %d %Y')`"
output: 
  html_document:
    keep_md: true
#  github_document
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating a HTTPS token

In order to be able to communicate with the git server, it is necessary to create a so-called Token. this token varify the identity of the person trying to connect to the server. 

Using R libraries such as `usethis` and `gitcreds` make it possible to do this from within R. 

```r
usethis::create_github_token()
```

will pop-up a new browser windows, where the user can create a new token.

It is always good to give the token a specific name, which will help identify, what it is for and why it was created. 

After the token was created, **remember to copy it first before continuing**. It is not possible to see the token again,s once the page was reloaded.

```r
gitcreds::gitcreds_set()
```

Without this token, one cannot communicate with the server, i.e. commiting a `pull` or `push` requests. 

The token created for this session is called "happygitwithr.com tutorial" and is valid for the next 30 days (expires on Sat, Jan 1 2022).


After that I will need to generate an ew token.

When the PAT expires, return to [https://github.com/settings/tokens](https://github.com/settings/tokens) and click on its Note. (You do label your tokens nicely by use case, right? Right?) At this point, you can optionally adjust scopes and then click “Regenerate token”. You can optionally modify its Expiration and then click “Regenerate token” (again). As before, **copy the PAT to the clipboard**, call `gitcreds::gitcreds_set()`, and paste!

### helpful command for token 

```r 
usethis::gh_token_help()

usethis::git_sitrep()
```

# Connect RsStudio to git and Github

Eventhough it is already connected, here aresome steps needed to be done to achieve this.

## Make a repo on github

First one need to create a repo on the [github site](https://github.com/). Initialize this repository with "Add a README file".

Copy a clone URL to your clipboard. If you’re taking our default advice, copy the HTTPS URL. But if you’re opting for SSH, then make sure to copy the SSH URL.

## Clone the repository to your computer via RStudio.

In RStudio, start a new Project:

- _File_ > _New Project_ > _Version Control_ > _Git_. In “Repository URL”, paste the URL of your new GitHub repository. It will be something like this.
- I suggest you check “Open in new session”, as that’s what you’ll usually do in real life.
- Click “Create Project”.

## Make local changes, save, commit

This is mainly just to check if it works, but really necessary, as we will do many changes.

From RStudio, modify the README.md file, e.g., by adding the line “This is a line from RStudio”. Save your changes.

Commit these changes to your local repo. How?

From RStudio:

- Click the “Git” tab in upper right pane.
- Check “Staged” box for README.md.
- If you’re not already in the Git pop-up, click “Commit”.
- Type a message in “Commit message”, such as “Commit from RStudio”.
- Click “Commit”.

## Push your local changes online to GitHub

Click the green “Push” button to send your local changes to GitHub.

## Confirm the local change propagated to the GitHub remote

Go back to the browser. I assume we’re still viewing your new GitHub repo.

Refresh.

done

## Clean up

remove the test repo and delete the local file.

## Solving possible problems

If you encounter any problems or even if you have found the appropriate solution - go [here](https://happygitwithr.com/troubleshooting.html).

# New projects

There are several ways to connect a project with github. 

- It can be a completely new project, which we first create a repository on Github and than a running it in RStudio.
- or it can be an existing project, which we first want to introduce it to a new gitHub repo.
- on the other hand it can be an RStudio project first, which we want to add to a github repo (so-called "github last").

## New project, GitHub first

This is the preferred method.

We create a new Project, with the preferred “GitHub first, then RStudio” sequence. Why do we prefer this? Because this method of copying the Project from GitHub to your computer also sets up the local Git repository for immediate pulling and pushing. Under the hood, we are doing `git clone`.

1. Make a repo on GitHub
2. New RStudio Project via git clone
    - _File_ > _New Project_ > _Version Control_ > _Git_. 
    - In the “repository URL” paste the URL of your new GitHub repository. It will be something like this https://github.com/yeroslaviz/myGit.git.
3. make changes and commit them
4. push to the repo
5. If changes were make on the github repository online (on Github), one needs to `pull` the changes to the local folder.

## Existing project, GitHub first

This is mainly repeating the steps from before for creating the repository on the github website and cloning it to a local folder. After cloning the repo one can just copy-paste the files from the existing projects into the newly created folder.

After the files are copied, one can marked then im the _staged_ box and `commit` -> `push`.

## Existing project, GitHub last

This an explicit workflow for connecting an existing local R project to GitHub, when for some reason you cannot or don’t want to do a “GitHub first” workflow.

When does this come up? 

Example: it’s an existing project that is already a Git repo with a history you care about. Then you have to do this properly.

Here we have several methods, depending on the fact if we're using the RStudio IDE and/or the `usethis` package. It is a straightforward process which can be confusing, but if done in the correct order should cause many problems. 

1. make sure you ahve a local project prepared with RStudio or create one using the `usethis::create_project()` command. 2. make and verify that you have a git repo with either the RStudio parameters or with the command `usethis::git_init()`.
3. stage and commit
4. connect the repo with Github.  This can be done with either the command `usethis::use_github()` or bith the help of the git command, after creating a repo online and copying the path to the repo.

```shell
git remote add origin https://github.com/jennybc/myrepo.git
git push --set-upstream origin main
```

5. refresh the browser to see, if all went well. 

If not, you can read a more detailed description [here](https://happygitwithr.com/existing-github-last.html).


# Rstudio and Markdown scripts

when using RStudio and git with a Github repository, it might be good to also take advantage of the Markdown possibilities of git. Ususally when rendering an HTML file, the `.Rmd` file is converted into a `.md` file which is then converted to HTML, deleting the md file in the process. 

It might be quite useful to keep this file.

This can be done in two ways:

1. either add the option `keep_md: true` to the title of the Rmarkdown file
2. or knit the file instead in a `output: github_document` instead of an html format. 

The big advantage of having an `.md` file is the ability of Github to present this file in an HTML-similar format. HTML files can be only read in the raw format, which is very user-unfriendly. 

If using the first option, it is always important to not forget also comiting and pushing the folder "FileName_files", which is created when the output file is html, but we still want to keep the `.md` file. This folder must also be uploaded to the Github repository in order for the Markdown file to be shown correctly, including the plots, pictures or images. 


```{r pressure, echo=FALSE, comment="test"}
plot(pressure)
```

The plot above is just an example of how the Markdown file creates the folder "gitwithR_files" with the plot image in it. 


If we are going to only create a script, we don't need all the overhead of the Rmarkdown, but we can create a `.R` file instead. This file can also be rendered into an HTML format, if needed. 

Onr only need to make a few changes to switch between an RMarkdown and an Rscript. 

```RMarkdown
Here's some prose in a very special comment. Let's summarize the built-in
dataset `VADeaths`.
here is a regular code comment, that will remain as such
 
'```{r}
summary(VADeaths)
'```

Here's some more prose. I can use usual markdown syntax to make things
**bold** or *italics*. Let's use an example from the `dotchart()` help to
make a Cleveland dot plot from the `VADeaths` data. I even bother to name
this chunk, so the resulting PNG has a decent name.

'```{r}
dotchart
dotchart(VADeaths, main = "Death Rates in Virginia - 1940")
'```
```


The above example is a standard RMarkdown format, while the one below is the way to modify an Rscript in order for it to be rendered. (the ticks `'` infront of the R code chunk are there so taht it will all be recognized as a comment).

- Anything that’s not R code? Like the YAML and the prose? Protect it with roxygen-style comments: start each line with `#'`.
- Anything that is R code? Let it exist “as is” as top-level code. That means you’ll need to change the syntax of R chunk headers like so:
    + Before: ` ```{r setup, include = FALSE} `
    + After: ` #+ r setup, include = FALSE `

```R
#' Here's some prose in a very special comment. Let's summarize the built-in
#' dataset `VADeaths`.
# here is a regular code comment, that will remain as such
summary(VADeaths)

#' Here's some more prose. I can use usual markdown syntax to make things
#' **bold** or *italics*. Let's use an example from the `dotchart()` help to
#' make a Cleveland dot plot from the `VADeaths` data. I even bother to name
#' this chunk, so the resulting PNG has a decent name.
#+ dotchart
dotchart(VADeaths, main = "Death Rates in Virginia - 1940")
```

The snippet above was copied to the script file `VADeaths.R` in order to test this.








# Session info

```{r}
R.version
sessionInfo()
```











